---
title: "Restaurant_Type"
author: "ACP"
date: "11/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Analyses for Boston restaurant violations final project exploring differences based on restaurant and restaurant type.

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
finalviol <- read.csv("finalviol.csv")
```

```{r}
finalviol <- finalviol %>%
  filter(year > 2006, LICSTATUS == "Active") %>%
  mutate(violstatus_fail = ifelse(ViolStatus == "Fail", 1, 0))

finalviol_anyfail <- finalviol %>%
  group_by(DESCRIPT, name, ZIP, LICENSENO, year) %>%
  summarize(fails = sum(violstatus_fail), num_cat = n()) %>%
  mutate(anyfails = ifelse(fails > 0, 1, 0), fivefails = ifelse(fails > 4, 1, 0), pcnt_fails = fails/num_cat)

finalviol_anyfail %>% group_by(DESCRIPT) %>%
  summarize(mean(anyfails), n())
```

There are four main categories of restaurants in the dataset:
  1. Eating & Drinking
  2. Eating & Drinking w/ Take Out
  3. Mobile Food Walk On
  4. Retail Food
  
Across all years, the vast majority of restaurants had at least one failure across the violation categories inspected (ranging from 98.1% among Retail Food to 99.7% among Mobile Food Walk On restaurants).

*Note: These analyses are limited to restaurants with active licenses only, but these numbers did not change substantially (fractions of percentage points) when restaurants with inactive licenses were also included.*

```{r}
finalviol_anyfail %>% group_by(DESCRIPT) %>%
  summarize(mean(fivefails), n())
```

Across all years, there was more variation across restaurant categories in the percent of restaurants per category that had at least five failures across violation categories. About a third of Eating and Drinking and Eating & Drinking w/ Take Out restaurants, about a fifth of Retail Food restaurants, and about a sixth of Mobile Food Walk On restuarants had at least five failures.

```{r}
finalviol_anyfail %>% filter(anyfails > 0) %>%
  group_by(DESCRIPT) %>%
  summarize(mean(pcnt_fails), n())
```

Among restaurants that failed at least one inspection category, Retail Food restaurants overall failed the highest percent of violation categories inspected with the average Retail Food restaurant failing 67.9% of the violation categories inspected. Eating & Drinking restaurants failed the lowest percent of categories inspected with the average Eating & Drinking restaurant failing 58.9% of violation categories inspected.

```{r}
finalviol_typeyear <- finalviol_anyfail %>%
  filter(anyfails > 0) %>%
  group_by(DESCRIPT, year) %>%
  summarize(avg_pcnt_fails = mean(pcnt_fails), count = n())

ggplot(finalviol_typeyear, aes(year, avg_pcnt_fails, color = DESCRIPT)) +
  geom_line() +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Percent of Violation Categories Failed", limits = c(0, 0.8))
```

With the exception of Mobile Food Walk On, all restuarants that failed at least one violation category inspected tended to fail a similar percent of violation categories inspected across years. This variation for Mobile Food Walk On may be driven by the low number of Mobile Food Walk On restaurants (ranging from 1 in 2009 to a peak of 102 in 2018).

```{r}
finalviol_cat <- finalviol %>%
  group_by(DESCRIPT, year, violcats) %>%
  summarize(avg_cat = mean(violstatus_fail), count = n())

ggplot(filter(finalviol_cat, DESCRIPT == "Eating & Drinking"), aes(year, avg_cat, color = violcats)) +
  geom_line() +
  ggtitle("Frequency of Violation Categories for Eating & Drinking Restuarants") +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Percent of Inspections Failed by Violation Category", limits = c(0, 0.6))
```

```{r}
ggplot(filter(finalviol_cat, DESCRIPT == "Eating & Drinking w/ Take Out"), aes(year, avg_cat, color = violcats)) +
  geom_line() +
  ggtitle("Frequency of Violation Categories for Eating & Drinking w/ Take Out Restuarants") +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Percent of Inspections Failed by Violation Category", limits = c(0, 0.6))
```

```{r}
ggplot(filter(finalviol_cat, DESCRIPT == "Mobile Food Walk On"), aes(year, avg_cat, color = violcats)) +
  geom_line() +
  ggtitle("Frequency of Violation Categories for Mobile Food Walk On Restuarants") +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Percent of Inspections Failed by Violation Category", limits = c(0, 1.0))
```

```{r}
ggplot(filter(finalviol_cat, DESCRIPT == "Retail Food"), aes(year, avg_cat, color = violcats)) +
  geom_line() +
  ggtitle("Frequency of Violation Categories for Retail Food Restuarants") +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Percent of Inspections Failed by Violation Category", limits = c(0, 0.7))
```

Across all restaurant types, the frequency of failing by violation category is approximately the same across violation categories across all years (generally within about a five percentage point band).

```{r}
finalviol_violcats <- finalviol %>%
  group_by(violcats, name, ZIP, LICENSENO, year) %>%
  summarize(fails = sum(violstatus_fail), num_cat = n()) %>%
  mutate(anyfails = ifelse(fails > 0, 1, 0), pcnt_fails = fails/num_cat)

finalviol_violcats %>% group_by(violcats) %>%
  summarize(mean(anyfails), n())
```

Across all violation categories, there are high rates of reatuarants failing at least one inspection

```{r}
finalviol_catyear <- finalviol_violcats %>%
  filter(anyfails > 0) %>%
  group_by(violcats, year) %>%
  summarize(avg_pcnt_fails = mean(pcnt_fails), count = n())

ggplot(finalviol_catyear, aes(year, avg_pcnt_fails, color = violcats)) +
  geom_line() +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Percent of Inspections Failed", limits = c(0, 0.7))
```

Among restaurants that failed in at least one category, there were approximately

```{r}
finalviol_repeat <- finalviol %>%
  group_by(name, DESCRIPT, ZIP, violcats) %>%
  summarize(mult_fails = sum(violstatus_fail)) %>%
  mutate(violrepeat = ifelse(mult_fails > 1, 1, 0))
```

Based on the data, the ten most common repeat offenders within the same violation category are:
  1. Subway (02115)
  2. Cosi (02110)
  3. Dunkin Donuts (02125)
  4. Dunkin Donuts (02128)
  5. Dunkin Donuts (02122)
  6. McDonalds (02135)
  7. U Food Grill (02128)
  8. Burger King (02128)
  9. Dunkin Donuts RMG (02128)
  10. Currito Burrito (02128)
  
```{r}
finalviol_DD <- finalviol %>%
  filter(name == "Dunkin Donuts" | name == "Dunkin" | name == "DUNKIN (STATION DONUTS)" | name == "DUNKIN DONUT" | name == "Dunkin Donuts (1447 Tremont St.)" | name == "Dunkin Donuts (209 North Harvard St.)" | name == "Dunkin Donuts (219 Cambridge)" | name == "Dunkin Donuts (48 W Broadway)" | name == "Dunkin Donuts (895 Morton St.)" | name == "DUNKIN DONUTS (ALLSTON)" | name == "DUNKIN DONUTS (BAY ROAD)" | name == "Dunkin Donuts (Canal St.)" | name == "DUNKIN DONUTS (CARNEY)" | name == "Dunkin Donuts (Causeway St.)" | name == "DUNKIN DONUTS (CITY PLACE)" | name == "Dunkin Donuts (Front Foodcourt)" | name == "Dunkin Donuts (TRMNL-C Baggage Area)" | name == "Dunkin Donuts (V.F.W. Pkwy.)" | name == "DUNKIN DONUTS (WEST ROXBURY)" | name == "DUNKIN DONUTS (WTC)" | name == "DUNKIN DONUTS @ HAYDEN HALL" | name == "Dunkin Donuts @ MBTA Station" | name == "Dunkin Donuts @ Stop & Shop" | name == "Dunkin Donuts at Old Colony" | name == "DUNKIN DONUTS BLUE HILL AVE" | name == "Dunkin Donuts Kiosk" | name == "Dunkin Donuts North Station" | name == "Dunkin Donuts Space No. 86" | name == "Dunkin Donuts-Amer Airlines/Landside" | name == "DUNKIN DONUTS-U S AIR" | name == "DUNKIN DONUTS'" | name == "Dunkin Donuts(757 Centre St.)" | name == "DUNKIN DONUTS(ARRIVAL)" | name == "DUNKIN DONUTS(DEPARTURE)" | name == "DUNKIN DONUTS(FRANKLIN)" | name == "DUNKIN DONUTS(SATELLITE)" | name == "DUNKIN DONUTS(WOLCOTT SQ. LLC)" | name == "Dunkin Donuts/1138 Washington St." | name == "Dunkin Donuts/Airside" | name == "DUNKIN DONUTS/BASKIN ROBBINS" | name == "DUNKIN DONUTS/BEE FINE FOODS" | name == "DUNKIN DONUTS/FIRST DONUTS INC" | name == "DUNKIN DONUTS/GALLERIA" | name == "DUNKIN DONUTS/GALLIVAN" | name == "DUNKIN DONUTS/HYDE PK AVE" | name == "DUNKIN DONUTS/PLAZA ENT" | name == "DUNKIN DONUTS/SCNVANOS" | name == "DUNKIN DONUTS/WALDWIN GROUP" | name == "DUNKIN'" | name == "Dunkin' Donuts" | name == "DUNKIN' DONUTS" | name == "Dunkin' Donuts Express" | name == "DUNKIN' DONUTS No. 0311" | name == "DUNKIN' DONUTS-WATERMARK" | name == "Dunkin' Donuts(C.A. Donuts)" | name == "DUNKIN' DONUTS(RMG DONUTS LLC)" | name == "DUNKIN/DONUT/WATERMARK") %>%
  group_by(LICENSENO, name, ZIP, year) %>%
  summarize(fails = sum(violstatus_fail), count = n()) %>%
  mutate(pcnt_fails = fails / count)

ggplot(finalviol_DD, aes(year, pcnt_fails, color = as.factor(LICENSENO))) +
  geom_line() +
  ggtitle("Dunkin Donuts Violations by Franchise per Year") +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Percent of Inspections Failed") +
  theme(legend.position = "none")
```

```{r}
ggplot(finalviol_DD, aes(year, pcnt_fails, group = year)) +
  geom_boxplot() +
  ggtitle("Dunkin Donuts Violations by Franchise per Year") +
  scale_x_continuous(name = "Year", limits = c(2006, 2019), breaks = c(2007:2018)) +
  scale_y_continuous(name = "Percent of Inspections Failed") +
  theme(legend.position = "none")
```

